---
output:
  pdf_document:
    # citation_package: biblatex
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
    template: ../svm-latex-ms.tex
title: "Political Donor Polarization in Wisconsin"
thanks: "Code and data available at: github.com/rossdahlke"
author:
- name: Ross Dahlke
affiliation: Stanford University
abstract: "American politics has recently been defined by unprecedented levels of partisan polarization. Given the concurrent rise of the amount of money in politics, there is a natural connection that many people make between money in politics and polarization. This paper uses the occurence of a specific event, former Wisconsin Governor Scott Walker's introduction and passage of Act 10 and subsequent protests and recall election, to analyze the relationship between donor polarization and mass polarization. The data show that political donor networks polarized during the 2012 election cycle at the same time as the mass electorate. This result suggests that politic donors were likely not the cause of the polarization in the state and provides evidence for the 'consumption' model of political donations."
keywords: "state politics, political donations, network analysis, polarization"
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 12pt
spacing: double
bibliography: ../bibliography.bib
biblio-style: apsr
---

Political campaign finance plays an important role in the American political system. This significance is evidenced by the attention that academic researchers pay to the topic as well as the many different contexts in which campaign finance is studied. For example, research has been conducted on the impact of political donations on roll-call voting in the U.S. Congress [@roscoe2005; @stratmann1991], gender representation in political parties [@crowder-meyer2018; @barber2016b; @kitchens2016; @thomsen2017], ability to win political campaigns [@bonica2017; @bonneau2007], the connection between money raised and public attention [@ellis2017], judicial function [@palmer2008], perceptions of corruption [@bowler2015], political economy and stock returns [@akey2015; @garro2020; @cooper2010], and the significant amount of time that candidates and legislators devote to fundraising [@torres-spelliscy2017]. 

Even though political donors are believed to play an out-sized role in democracy, the psychological processes of donors is thought to be similar to voters. Political donations can be thought of an extension of voting. In other words, both actions are political consumptions that seek to improve their preferred candidate's chances of winning. Ansolabehere, de Figueiredo and Snyder summarized this idea by stating, "In our view, campaign contributing should not be viewed as an investment, but rather as a form of consumption--or, in the language of politics, participation" [@ansolabehere2003]. Donations can be seen as an outlet for motivated citizens to increase their participation beyond just turning out to vote when they perceive the stakes of elections to be high [@hill2017].

The folk-theory of political donors is of smokey backrooms and access-oriented donors who seek to have a direct influence on policy making. However, even when donors contribute to legislators that maximize their economic interests, donations are not found to be motivated by existing policy agreements and not an expectation of access [@barber2016c]. Even donations from business executives have been found to be "best understood as purchases of 'good will' whose returns, while positive in expectation, are contingent and rare" [@gordon2007].

Although the psychological process of making a political campaign contribution can be thought of as similar to voting, there are significant demographic and ideological differences between donors and voters. People with lower incomes, less education, and do not work in professional and managerial jobs are less likely to be politically engaged, including making political donations [@laurison2016]. Donors to the Democratic and Republican parties were summarized as being "Limousine Liberals" and "Corporate Conservatives" [@francia2005].In addition, Democrats and Republicans draw their bases of electoral support from different geographic bases. However, major campaign donors are highly concentrated geographically. These "big-donor neighborhoods" are unrepresentative of the country as a whole and point to these communities having a distinct political culture [@bramlett2011]. In both parties, donors are more ideologically extreme than non-donating voters, even among primary voters [@hill2017; @francia2003]. Even wealthy donors who make up the "big money" in politics are very partisan [@mccarty2006]. 

This ideological extremity shown by political donors has led some scholars to suggest that political donors are contributors to the partisan polarization of the politics of the United States [@francia2005]. This speculation that that political donors are contributors of polarization is bolstered by the observation that both political polarization and campaign spending have risen in conjunction [@mccarty2006]. However, there little evidence for the causal relationship of donors causing political polarization. And many have concluded that political donations don't influence polarization [@harden2016; @laraja2011; @keena2019]. Furthermore, it is likely the case that the causal arrow flows the other direction, and it is more polarized candidates and electorate that have led to more polarized donors [@harden2016; @laraja2011; @keena2019].

Studying polarization, particularly among political donors can be difficult because of the myriad of potential confounding factors that can contribute to polarization [@harden2016]. In addition, polarization is generally a phenomenon that gradually increases or decreases over time [@pew2017]. However, this paper leverages a singular event, former Wisconsin Governor Scott Walker's proposition and passage of Act 10, a "budget repair bill" that ended collective bargaining for teachers unions and the subsequent protests and recall election, to examine political donor polarization in the state of Wisconsin. Given the recent research which has pointed to the polarization of political donors as being _reactive_ instead of _causal_ to broader polarization, we could expect political donors to follow the trend of voters and polarize after introduction of Act 10 and subsequent events.

**H1: Political donors in the State of Wisconsin polarized during the 2011-2012 election cycle compared to the 2009-2010 election cycle and maintained their level of polarization in the 2013-2014 election cycle.**

If hypothesis one is found to be correct, these results would strengthen the evidence for politics donors being _reactive_ to their political environment as we would expect under Ansolabehere, de Figueiredo and Snyder's consumption model of political giving.

Alternatively, if political donors are contributors to polarization, as is suggested by some scholars, we would expect to see hypothesis 2. 

**H2: Polarization levels stays the same from 2009-2010 compared to 2011-2012.**

If hypothesis two is upheld, the result would suggest that political donors helped to create the polarized political environment that we see today.

In addition, this paper makes a methodological contribution to the field of political polarization studies by taking a network approach to measuring polarization to similar to studies of congressional polarization [@waugh2009, @zhang2008] and polarization in social media networks [@guerra2013; @garcia2015; @conover2011] and uses modularity as a measure of polarization within political donor networks. This paper conceives of the political donor landscape of donors and candidates acting as nodes who are connected by donations made that act as edges. This method is important in studying political donor networks because it takes into consideration real-world actions, such as in the study of polarization among member of congress where voting records [@guerra2013] and co-sponsorships [@zhang2008], which are used to study polarization opposed to surveys administered to donors that rely on self-reported ideology and partisanship.

# Wisconsin Context

Both Wisconsin's legislators and mass public are among the most polarized in the nation [@cramer2016] and has been used by academics as an example of how political actions change in contenious and divisive environments [@bode2018]. Although many state legislatures are also experiencing polarization [@shor2015], Wisconsin is unique in that there is a single event that many point to in creating "the most politically divisive place in America" [@kaufman2012]. 

In 2011, newly-elected Republican Governor Scott Walker introduced Act 10, a "budget reconciliation bill" that among other cuts to funding, stripped public school teachers of collective bargaining via their union. Up to 100,000 protested this "anti-union bill" at the State Capitol and even occupied the capitol building for a period of time [@sewell2011]. Democratic lawmakers fled to Illinois in an effort to delay or stop the bill from passing into law [@layton2011]. In 2012 there was an unsuccessful election to recall Governor Walker.

Wisconsin Governor Scott Walker's self-anointed "divide and conquer" politics [@blake2012] has left a political divide in Wisconsin that persists to today. The result is that "divisive politics ruled Wisconsin over the last decade" [@marley2019]. The Marquette Law School poll headed by Charles Franklin has called public opinion in Wisconsin a "lesson in the two worlds of Wisconsin" where "it seems often as if people have not only differing opinions but differing views of facts and realities" [@borsuk2017].

```{r setup, include=FALSE}
knitr::opts_chunk$set(include = FALSE)
library(tidyverse)
library(infer)
library(igraph)
```

# Methodology

  All data on political contributions came from the Wisconsin Campaign Finance Information System (CFIS). I exported all contributions to State Assembly, State Senate, and Gubernatorial races from the 2010, 2012, and 2014 elections. This dataset does not include donations to party committees, although it does include disbursements from these committees. I manually created a table of the parties of each of all the campaigns receiving contributions in this timeframe and added the party of the campaign receiving the donation to this dataset.

```{r}
donations <- readRDS("../data/wi_donations.RDA") %>% 
  mutate(source = str_to_lower(str_replace(source, "[[:punct:]]", " ")),
         election_year =  as.character(election_year)) 

donations_n <- donations %>% 
  count() %>% 
  pull()

anon_donations_n <- donations %>% 
  filter(str_detect(source, c("unitemized|anonymous")) == T) %>% 
  count() %>% 
  pull()

donations_2 <- donations %>% 
  filter(str_detect(source, c("unitemized|anonymous")) == F) %>% 
  mutate(zip_5 = str_sub(zip, 1, 5),
         refined_source = refinr::n_gram_merge(refinr::key_collision_merge(source)),
         refined_source_zip = paste0(refined_source," : ",zip_5)) 

filtered_donations <- donations_2 %>% 
  group_by(election_year, refined_source_zip) %>% 
  mutate(source_count = n()) %>% 
  filter(source_count > 1) %>% 
  group_by(election_year, target) %>% 
  mutate(target_count = n()) %>% 
  filter(target_count > 20) %>% 
  ungroup()
```

I started with `r scales::comma(donations_n)` donations. To clean the data, I filtered out `r scales::comma(anon_donations_n)` unitemized/ anonymous donations , removed punctuation from the names of the donors, and used Open Refine via the `refiner` R package to standardize names (for example, Jim versus James). Next, I created a unique identifier for donors by combining their standardized name with their zip code. This identifier was created to be able to link donors who contributed across multiple campaigns in multiple years without considering two different people, with the same name, from different locations to be the same person.

Next, I derived the partisanship of each donor in each election cycle. I calculated each donor's partisanship by taking the percent of donations that each donor gave to Republicans divided by their donations to Republicans and Democrats. I took that "percent donated to Republicans" and rescaled it from -1 to 1, where -1 represents the most Democratic donors, and 1 the most Republican donors. I also calculated each individual's party bin: if more than 75% of donations were to Democrats, they were labeled as a Democrat; if more than 75% of donations were to Republicans, they were labeled as a Republican; if their donations were somewhere inbetween, they were labeled as being a bipartisan donor.

```{r}
by_donor <- filtered_donations %>% 
  filter(party != "other") %>% 
  group_by(election_year, refined_source_zip, party) %>% 
  summarize(contribution = sum(contribution)) %>% 
  ungroup() %>% 
  pivot_wider(names_from = party, 
              values_from = contribution) %>% 
  mutate(rep = replace_na(rep, 0),
         dem = replace_na(dem, 0),
         total_contributions = rep + dem,
         per_rep = rep / total_contributions,
         partisanship = scales::rescale(per_rep, to = c(-1, 1)),
         abs_partisanship = abs(partisanship),
         party_bin = case_when(
           per_rep < .25 ~ "democrat",
           per_rep >= .25 & per_rep <= .75 ~ "bipartisan",
           per_rep > .75 ~ "republican"
         ))

```

To quantify the levels of polarization in each election cycle, I calculated two statistics: network modularity and average absolute partisanship of donors.

First, political donations can be thought of as a network where donors and candidates are nodes and donations connecting donors and candidates are edges. This conceptualization of the political donor landscape as network allows us to examine the network structure and calculate network statistics on the graph of donors and candidates. One of the most useful network statistics for measuring polarization in a network's modularity.

The modularity of a graph measures how good the division of groups (such as political parties) is by calculating "the number of edges falling within groups minus the expected number in an equivalent network with edges placed at random" [@newman2006]. The modularity of a network falls in range [-1/2, 1]. If the modularity is positive, the number of edges that remain within each group is greater than the expected number to remain in-group based on chance. The higher the modularity, the greater the concentration of edges within each groups. In other words, the higher the modularity of a network, the higher the polarization among the groups. Formally, the equation to calculation modularity Q is:

$$Q = \frac{1}{2m} \sum_{ij}\left[A_{ij} - \frac{k_{i}k_{j}}{2m} \right]\delta(g_{i},g_{j})$$

I calculated the modularity of the network graphs of each election cycle (2010, 2012, 2014) using the `igraph` R package [@igraph]. I used candidates' declared parties and donors' party bin as the groups for the modularity calculation. The modularity of the network graph of each election is in Table 1.

```{r}
modularity_calc <- function(year = year){
  nodes_w_party <- by_donor %>% 
  filter(election_year == year) %>% 
  select(refined_source_zip, party_bin) %>% 
  rename(node = refined_source_zip,
         party = party_bin) %>% 
  rbind(donations_2 %>% 
  distinct(target, party) %>% 
  filter(party != "other") %>% 
    rename(node = target) %>% 
    mutate(party  = case_when(
      party == "rep" ~ "republican",
      party == "dem" ~ "democrat"
    ))) %>% 
  distinct(node, party) %>% 
  mutate(party_num = case_when(
    party == "democrat" ~ 1,
    party == "bipartisan" ~ 2,
    party == "republican" ~ 3
  ))
  

filtered_donations %>%
  filter(election_year == year & party != "other") %>% 
  select(refined_source_zip, target, contribution) %>% 
  graph_from_data_frame(vertices = nodes_w_party) %>% 
  modularity(nodes_w_party$party_num)
}

```

```{r}
election_years <- by_donor %>% 
  distinct(election_year) %>% 
  pull()
```

```{r}
tibble(election_years) %>% 
  group_by(election_years) %>% 
  mutate(modularity = modularity_calc(election_years))
```

In addition to calculating the change in modularity of each of the election cycles, I also analyzed the change in mean absolute partisanship of the donors in each election cycle. 

I defined a donor's absolute partisanship as the absolute value of their partisanship score (which is on a scale from -1 to 1). Therefore, the larger a donor's absolute the partisanship, the higher percentage of their money that they contributed to a single party. To calculate the significance in the difference of the mean absolute partisanship, I use a bootstrap methodology with 1000 replications. The results of the bootstrap are found in Table 2.

```{r}
diff_boostrap <- function(year_1 = "2014",
                          year_2 = "2012",
                          replications = 100) {
  year_1 <- as.character(year_1)
  year_2 <- as.character(year_2)
  
  bootstrap <- by_donor %>% 
    filter(election_year %in% c(year_1, year_2)) %>% 
    specify(abs_partisanship ~ election_year) %>% 
    generate(reps = replications, type = "bootstrap") %>% 
    calculate(stat = "diff in means", order = c(year_1, year_2))
  

  bootstrap %>% 
    get_ci() %>% 
    cbind(bootstrap %>% 
            summarize(mean_diff = mean(stat))) %>% 
    cbind(bootstrap %>% 
            get_p_value(obs_stat = 0, direction = "two_sided"))
  

                          }
```


```{r}
election_years_expanded <- expand.grid(year1 = as.character(election_years),
            year2 = as.character(election_years)) %>% 
  filter(year1 != year2 & 
           as.numeric(year1) > as.numeric(year2) & 
           (as.numeric(year1) - as.numeric(year2) < 2)) %>% 
  ungroup() 

partisanship_diff_results <- cbind(election_years_expanded, 
                                   Map(diff_boostrap, 
                                       election_years_expanded$year1, 
                                       election_years_expanded$year2, 
                                       1000) %>% 
                                     tibble() %>% 
                                     unnest())


```

```{r}
ggplot(by_donor %>% filter(abs_partisanship < 1), aes(abs_partisanship, fill = party_bin)) +
  geom_histogram() +
  facet_wrap(party_bin~election_year, scales = "free_y")
```

```{r}
ggplot(by_donor %>% filter(abs_partisanship < 1), aes(abs_partisanship, fill = party_bin)) +
  geom_density() +
  facet_wrap(party_bin~election_year)
```


# References
